Bipartite Fidelity and Loschmidt Echo of Bosonic Conformal Interface

### 2017-12-01 analy_eval.tex
[paper] Proofread very quickly...

In this subsection, we relate the free energy to the amplitudes between boundary states and present the analytic results.

@@

One subtlety of the above description is that the two semi-circles in the center diagrams of Fig.~\ref{fig:fidel-map} and Fig.~\ref{fig:H-tau_fold} are not precisely concentric. This can be resolved by the following observation. There exists a conformal map $\zeta(\xi)$ that maps the non-concentric circles to the two standard concentric circles of radii $1$ and $R$ ($R>1$) on the $\zeta$ plane\cite{brown_complex_2009}. Then the logarithmic map $w = \ln \zeta$ produces a cylinder of width $\ln R$. In our case, since the height of the cylinder is always $\pi$, the width of the cylinder is a conformal invariant that only depends on the cross ratio of the half annulus. The four intersection points of two standard concentric circles on $\zeta$ plane are $(\pm 1,0)$ and $(\pm R,0)$, whose cross ratio is
\begin{equation}
\eta = \frac{(1 + R)^2}{(1 - R)^2}. 
\end{equation}
Hence the width of the cylinder is $\ln \frac{\sqrt{ \eta } - 1}{\sqrt{ \eta} + 1}$. Since conformal transformation preserves the cross ratio, the result is the same if we use the cross ratio of the slightly non-concentric diagrams of Fig.~\ref{fig:fidel-map} and Fig.~\ref{fig:H-tau_fold}. The calculation in Eq.~\eqref{eq:fidel_cyd_width} and Eq.~\eqref{eq:echo_cyd_width} equivalently use the leading order approximation to $\eta$ in the respective geometries and thus get the leading order term in the width of the cylinders. The slight deviation to the precise concentric geometry will only bring in $\frac{\epsilon}{L}, \frac{\epsilon}{\tau}$ corrections to $\eta$ and the width parameter, so will not affect the fidelity and echo exponents.

>>>>

In this subsection, we relate the free energy to the amplitudes between boundary states, and present the analytic results.

@@

One subtlety of the above description is that the two semi-circles in the center diagrams of Fig.~\ref{fig:fidel-map} and Fig.~\ref{fig:H-tau_fold} are not precisely concentric. This can be resolved by the following observation. There exists a conformal map $\zeta(\xi)$ that maps the non-concentric circles to the two standard concentric circles of radii $1$ and $R$ ($R>1$) on the $\zeta$ plane\cite{brown_complex_2009}. Then the logarithmic map $w = \ln \zeta$ produces a cylinder of width $\ln R$. In our case, since the height of the cylinder is always $\pi$, the width of the cylinder is a conformal invariant that only depends on the cross ratio of the half annulus. The four intersection points of two standard concentric circles on $\zeta$ plane are $(\pm 1,0)$ and $(\pm R,0)$, whose cross ratio is
\begin{equation}
\eta = \frac{(1 + R)^2}{(1 - R)^2}. 
\end{equation}
Hence the width of the cylinder is $\ln \frac{\sqrt{ \eta } - 1}{\sqrt{ \eta} + 1}$. Since conformal transformation preserves the cross ratio, the result is the same if we use the cross ratio of the slightly non-concentric diagrams of Fig.~\ref{fig:fidel-map} and Fig.~\ref{fig:H-tau_fold}. The calculation in Eq.~\eqref{eq:fidel_cyd_width} and Eq.~\eqref{eq:echo_cyd_width} equivalently use the leading order approximation to $\eta$ in the respective geometries and thus get the leading order term in the width of the cylinders. The slight deviation to the precise concentric geometry will only bring in $\frac{\epsilon}{L}, \frac{\epsilon}{\tau}$ corrections to $\eta$ and the width parameter, which will not affect the fidelity and echo exponents.

### 2017-12-01 def_fidel_echo.tex
[paper] referee's suggestions done.
[paper] Proofread very quickly...

Fidelity is the square of the overlap of the groundstates of the two Hamiltonians, 
\begin{equation}
{\rm fidelity} \equiv |\langle \psi_1 |\psi_2  \rangle |^2 .
\end{equation}
For the systems we considered, $|\psi_1 \rangle$ is the groundstate of the two disconnected chains (of equal length $L$, hence ``bipartite") and $|\psi_2\rangle$ is that of the connected chains with conformal interface. Both of them can be produced by an imaginary time evolution. Taking the horizontal axis as imaginary time direction, the fidelity can be diagrammatically represented in Fig.~\ref{fig:fidel}, where the slits represent the disconnected boundary conditions, such as Dirichlet(D) and Neumann(N), and the dashed line represents the conformal interface parameterized by $\lambda$. The logarithmic fidelity is then (twice) the free energy of this diagram
\begin{equation}
\mathcal{F}( {\rm fidelity} )  = - \ln \langle \psi_1 |\psi_2 \rangle^2 = - 2 \ln |Z|.
\end{equation}

@@

The Loschmidt echo is also the (square of the) overlap of two wavefunctions. One of them is the groundstate of the disconnected chains and the other is the groundstate evolved by the Hamiltonian of the connected chains
\begin{equation}
\mathcal{L}( t )  \equiv |\langle \psi_{\rm gnd}  | e^{-i H t } | \psi_{\rm gnd} \rangle|^2 .
\end{equation}
The imaginary time version $\mathcal{L}( \tau  ) = |\langle \psi_{\rm gnd}  | e^{-H \tau } | \psi_{\rm gnd} \rangle|^2$ has a path integral definition similar to Fig.~\ref{fig:cut-and-join}, but to be consistent with the fidelity diagram, we take the horizontal axis as imaginary time and present it in Fig.~\ref{fig:echo}. Viewing the diagram as a partition function subject to the switching of boundary conditions, the logarithmic Loschmidt echo is also the associated free energy. After obtaining the free energy in imaginary time, we can analytically continue back to real time to get the $t$ dependence. For simplicity and comparison with the fidelity result, we will take the length of both of the chains to be $L$ and set $L \gg t$, leaving $t$ the only length scale in the echo calculation (Fig.~\ref{fig:H-tau_fold}). The dependence on nonzero $\frac{t}{L}$ and the asymmetry of the lengths of the chains will not be discussed here (see the treatment in Ref.~\onlinecite{stephan_local_2011} for special values of $\lambda$).

@@

If the interface is completely transparent, i.e. at the special point of $\lambda = 1$, the tip of the slit can be regarded as a corner singularity. According to Cardy and Peschel\cite{cardy_finite-size_1988}, the singularity will contribute to a term that is logarithmic of the corner's characteristic size, which is $\ln L$ in the fidelity and $\ln \tau$ in the Loschmidt echo. One would expect the fidelity and echo to have power law decay with respect to these scales in the long wavelength limit. In fact, the computations have been done in Ref.~\onlinecite{stephan_logarithmic_2013,stephan_local_2011,vasseur_universal_2014,vasseur_crossover_2013,kennes_universal_2014} using either the Cardy-Peschel formula or the integration version of the Ward identity. If the slits boundary conditions are taken to be Dirichlet, we have the universal behavior for the leading term \cite{stephan_logarithmic_2013,stephan_local_2011}
\begin{equation}
\label{eq:full-pass}
\begin{aligned}
  \mathcal{F}({\rm fidelity}) &=  \frac{c}{8} \ln L, \\
 \mathcal{F}( {\rm echo} )  &= \frac{c}{4} \ln |\tau| \rightarrow \frac{c}{4} \ln |it + \epsilon|   \sim \frac{c}{4}\ln t ,
\end{aligned}
\end{equation}
where we have performed analytic continuation $\tau \rightarrow it + \epsilon$ with $\epsilon \rightarrow 0$ for echo.

>>>>

Fidelity is the square of the overlap of the groundstates of the two Hamiltonians, 
\begin{equation}
{\rm fidelity} \equiv |\langle \psi_1 |\psi_2  \rangle |^2 .
\end{equation}
For the systems we considered, $|\psi_1 \rangle$ is the groundstate of the two disconnected chains (of equal length $L$, hence ``bipartite") and $|\psi_2\rangle$ is that of the connected chains with conformal interface. Both of them can be produced by an imaginary time evolution. Taking the horizontal axis as imaginary time direction, the fidelity can be diagrammatically represented in Fig.~\ref{fig:fidel}, where the slits represent the disconnected boundary conditions, such as Dirichlet(D) and Neumann(N), and the dashed line represents the conformal interface parameterized by $\lambda$. The logarithmic fidelity is then (twice of) the free energy of this diagram
\begin{equation}
\mathcal{F}( {\rm fidelity} )  = - \ln \langle \psi_1 |\psi_2 \rangle^2 = - 2 \ln |Z|.
\end{equation}

@@

The Loschmidt echo is also (square of) the overlap of the two wavefunctions. One of them is the groundstate of the disconnected chains and the other is the groundstate evolved by the Hamiltonian of the connected chains
\begin{equation}
\mathcal{L}( t )  \equiv |\langle \psi_{\rm gnd}  | e^{-i H t } | \psi_{\rm gnd} \rangle|^2 .
\end{equation}
The imaginary time version $\mathcal{L}( \tau  ) = |\langle \psi_{\rm gnd}  | e^{-H \tau } | \psi_{\rm gnd} \rangle|^2$ has a path integral definition similar to Fig.~\ref{fig:cut-and-join}, but to be consistent with the fidelity diagram, we take the horizontal axis as imaginary time and present it in Fig.~\ref{fig:echo}. Viewing the diagram as a partition function subject to the switching of boundary conditions, the logarithmic Loschmidt echo is also the associated free energy. After obtaining the free energy in imaginary time, we can analytically continue back to real time to get the $t$ dependence. For simplicity and comparison with the fidelity result, we will take the length of both of the chains to be $L$ and set $L \gg t$, leaving $t$ the only length scale in the echo calculation (Fig.~\ref{fig:H-tau_fold}). The dependence on nonzero $\frac{t}{L}$ and the asymmetry of the lengths of the chains will not be discussed here (see the treatment in Ref.~\onlinecite{stephan_local_2011} for special values of $\lambda$).

@@

If the interface is completely transparent, i.e. at the special point of $\lambda = 1$, the tip of the slit can be regarded as a corner singularity. According to Cardy and Peschel\cite{cardy_finite-size_1988}, the singularity will contribute a term that is logarithmic of the corner's characteristic size, which is $\ln L$ in the fidelity and $\ln \tau$ in the Loschmidt echo. One would expect the fidelity and echo to have power law decay with respect to these scales in the long wavelength limit. In fact, the computations have been done in Ref.~\onlinecite{stephan_logarithmic_2013,stephan_local_2011,vasseur_universal_2014,vasseur_crossover_2013,kennes_universal_2014} using either the Cardy-Peschel formula or the integral version of the Ward identity. If the slits boundary conditions are taken to be Dirichlet, we have the universal behavior for the leading term \cite{stephan_logarithmic_2013,stephan_local_2011}
\begin{equation}
\label{eq:full-pass}
\begin{aligned}
  \mathcal{F}({\rm fidelity}) &=  \frac{c}{8} \ln L, \\
 \mathcal{F}( {\rm echo} )  &= \frac{c}{4} \ln |\tau| \rightarrow \frac{c}{4} \ln |it + \epsilon|   \sim \frac{c}{4}\ln t ,
\end{aligned}
\end{equation}
where we have performed analytic continuation $\tau \rightarrow it + \epsilon$ with $\epsilon \rightarrow 0$ for echo.

### 2017-12-01 free_boson_lattice.tex
[paper] Proofread very quickly...

We consider two harmonic chains with bosonic field $\phi_i$ and conjugate momentum $\pi_i$ at lattice site $i$. The left and right chains are connected between site $0$ and $1$ with the Hamiltonian
\begin{equation}
\begin{aligned}
\label{eq:lattice_H}
H &= \frac{1}{2} \sum_i \pi_i^2  +  \frac{1}{2} \sum_{i\ne 0 }  ( \phi_i - \phi_{i+1} )^2 \\
\quad & + \frac{1}{2} \begin{pmatrix}  \phi_0, \phi_1 \end{pmatrix}
\begin{bmatrix}
1 + \Sigma_{11}  & \Sigma_{12} \\
\Sigma_{21} &  1 + \Sigma_{22} \\
\end{bmatrix}
\begin{pmatrix}
\phi_0 \\
\phi_1 
\end{pmatrix},
\end{aligned}
\end{equation}
where the $2\times2$ matrix $\Sigma$ parameterizes the two-site interaction. We performed the standard scattering analysis in App.~\ref{app:interface_free_boson}. For mode with momentum $k$, the only possible scale invariant $S$-matrix is
\begin{equation}
S = - e^{ika} \Sigma,
\end{equation}
where $a$ is the lattice constant. In the continuum limit $a\rightarrow0$, the matrix $\Sigma$ can be parameterized as
\begin{equation}
\label{eq:def_Sigma}
\Sigma = -\lim_{a \rightarrow 0 } S = 
\begin{bmatrix}
\frac{\lambda^2- 1}{1 + \lambda^2} & \frac{-2\lambda }{1 + \lambda^2} \\
\frac{-2\lambda }{1 + \lambda^2} & \frac{1- \lambda^2}{1 + \lambda^2} \\
\end{bmatrix},
\end{equation}
where $\lambda\in\mathbb{R}$.

>>>>

We consider two harmonic chains with bosonic field $\phi_i$ and conjugate momentum $\pi_i$ at the lattice site $i$. The left and right chains are connected between site $0$ and $1$ with the Hamiltonian
\begin{equation}
\begin{aligned}
\label{eq:lattice_H}
H &= \frac{1}{2} \sum_i \pi_i^2  +  \frac{1}{2} \sum_{i\ne 0 }  ( \phi_i - \phi_{i+1} )^2 \\
\quad & + \frac{1}{2} \begin{pmatrix}  \phi_0, \phi_1 \end{pmatrix}
\begin{bmatrix}
1 + \Sigma_{11}  & \Sigma_{12} \\
\Sigma_{21} &  1 + \Sigma_{22} \\
\end{bmatrix}
\begin{pmatrix}
\phi_0 \\
\phi_1 
\end{pmatrix},
\end{aligned}
\end{equation}
where the $2\times2$ matrix $\Sigma$ parameterizes the two-site interaction. We performed the standard scattering analysis in App.~\ref{app:interface_free_boson}. For modes with momentum $k$, the only possible scale invariant $S$-matrix is
\begin{equation}
S = - e^{ika} \Sigma,
\end{equation}
where $a$ is the lattice constant. In the continuum limit $a\rightarrow0$, the matrix $\Sigma$ can be parameterized as
\begin{equation}
\label{eq:def_Sigma}
\Sigma = -\lim_{a \rightarrow 0 } S = 
\begin{bmatrix}
\frac{\lambda^2- 1}{1 + \lambda^2} & \frac{-2\lambda }{1 + \lambda^2} \\
\frac{-2\lambda }{1 + \lambda^2} & \frac{1- \lambda^2}{1 + \lambda^2} \\
\end{bmatrix},
\end{equation}
where $\lambda\in\mathbb{R}$.
